Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to boundary detection tasks

### 2023-07-02 discussion/discussion.tex
Updates from Overleaf

In this work, we introduced a method that leverages self-supervised deep neural networks to learn low-dimensional music latent representations with applications to music boundary detection. Building on existing approaches and architectures, we replaced traditional features with deep embeddings trained to represent high-level musical content information to analyze its performance in music segmentation tasks.

While not outperforming state-of-the-art existing baselines in contrast to similar research findings such \cite{deepfeaturesegment, SalamonDeepSegmentation}, our musically-informed technique has significant potential for boundary detection tasks given its overall performance improvement alongside the training set enlargement (GTZAN compared to Million Song Dataset). Most likely, so does for nearly all MIR downstream tasks that are not purely sonic-based.

>>>>

In this work, we introduced a method that leverages self-supervised deep neural networks to learn low-dimensional music latent representations with applications to music boundary detection. Building on existing approaches and architectures, we replaced traditional features with deep embeddings trained to represent high-level musical information to analyze its performance in music segmentation tasks.

While not outperforming state-of-the-art existing baselines and in contrast to similar research findings such \cite{deepfeaturesegment, SalamonDeepSegmentation}, our musically-informed technique has significant potential for boundary detection tasks given its overall performance improvement alongside the training set enlargement. Most likely, so does for nearly all MIR downstream tasks that are not purely sonic-based.

### 2023-07-02 future work/future work.tex
Updates from Overleaf

\begin{itemize}
  \item Transforms
  \item Stems
  \item Different takes
  \item k-fold cross validation
  \item Hyper-params: kernel size $0.005$ times the sample rate
  \item Larger representation layer
  \item Easy triplet mining \cite{XuanImprovedMining}
  \item Different boundary detection algorithms \cite{sf} was used because it was the default in MSAF \cite{MSAF}
\end{itemize}

Loss function margin. A margin of 0.2 can be a starting point, but it might not be the best value for every task. It's recommended to experiment with different values and choose the one that provides the best performance on your validation set.

Utilize the dB-scale Mel-spectrum magnitude and CQT of audio as input data; it is a popular choice for input representation in music-related tasks when applying CNNs, as has been demonstrated in various studies \cite{Kim2020OneStrategies}. While raw audio is the most meaningful audio representation, it's been reported that dB-scale Mel-spectrum frequency-domain summarization, grounded in psycho-acoustics, is computationally efficient and challenging to replicate solely through data-driven methods \cite{Kim2020OneStrategies}; therefore, a trade-off that looks worthy of being explored.

Visual and listening evaluation: 2D or 3D latent space visualization as per \ref{fig:manifold}. Arranging the embedding space in a visual display to evaluate the clustering of this sophisticated musical content to see to what extent they consider sonic attributes.

>>>>

This research merely scratches the surface, and vast territories are yet uncharted. In future studies, the following areas will be explored to extend the current research further:

\begin{itemize}
  \item Investigate the effect of different transformations and augmentation pipelines. Incorporating track stems and different takes of the same piece as natural, human-made augmentations could yield interesting results.
  \item Expand input data to include dB-scale Mel-spectrum magnitude and CQT of audio. This approach has been widely used in music-related tasks with CNNs \cite{Kim2020OneStrategies}. Though raw audio provides a rich representation, dB-scale Mel-spectrum offers a frequency-domain summarization that is not only grounded in psycho-acoustics but is also computationally efficient and hard to reproduce solely through data-driven methods \cite{Kim2020OneStrategies}. Therefore, this trade-off is worth exploring.
  \item Implement k-fold cross-validation to improve the robustness of the model's performance.
  \item Experiment with different hyperparameters. For instance, setting the kernel size to $0.005$ times the sample rate could match the Just Noticeable Difference (JND). As for the loss function margin, a starting point of 0.2 has proven effective, but different values should be explored to optimize performance on the validation set.
  \item Increase the size of the representation layer to 512 or 1024 dimensions.
  \item Apply easy triplet mining to improve the model's performance \cite{XuanImprovedMining}.
  \item Implement visual and auditory evaluations: 2D or 3D visualization of the latent space \ref{fig:manifold}, coupled with a Graphical User Interface (GUI) that enables playback for evaluation, can help assess the extent to which the model considers sonic attributes. It would also facilitate the understanding of the clustering of complex musical content. 
\end{itemize}

### 2023-07-02 methods/methods.tex
Updates from Overleaf

\section{Materials}
\input{methods/subsections/model}

\section{Convolutional Neural Network (CNN)}
\input{methods/subsections/test}
\newpage

>>>>

This work concentrates on employing unsupervised training of neural networks to extract meaningful features for music segmentation methodologies. As assertively stated by \cite{deepfeaturesegment}, this constitutes a logical evolution in the application of contemporary machine learning techniques to the issue of music segmentation.

Our work confronts the persistent challenge of acquiring labeled data for music segmentation, an endeavor often time-consuming and expensive. By capitalizing on the vast volumes of unlabeled music data, we bypass the need for laborious hand annotation, thereby offering a pathway to potentially more comprehensive generalization given the broader dataset.

\input{methods/subsections/deep learning}
\input{methods/subsections/implementation}

\newpage

### 2023-07-02 results/subsections/tables.tex
Updates from Overleaf

The GitHub repository containing all the code needed to run the experiments can be found \href{https://github.com/oriolcolomefont/Master-Thesis.git}{HERE}.The results are displayed in \ref{ta:results}

@@

\newsavebox\mfccSF
\begin{lrbox}{\mfccSF}
\centering
   $\begin{aligned}
     0.173 & \quad 0.185 & \quad 0.172 % SF algorithm
    \end{aligned} $
\end{lrbox}

\newsavebox\mfccFoote
\begin{lrbox}{\mfccFoote}
\centering
   $\begin{aligned}
     ... & \quad ... & \quad ... % Foote algorithm
    \end{aligned} $
\end{lrbox}

\newsavebox\mfccCNMF
\begin{lrbox}{\mfccCNMF}
\centering
   $\begin{aligned}
     ... & \quad ... & \quad ... % CNMF algorithm
    \end{aligned} $
\end{lrbox}

@@

\begin{table}
  \centering
  \begin{adjustbox}{width=\textwidth}
  \begin{threeparttable}
    \begin{tabular}{c|l|c|c|c} 
\toprule
     \multicolumn{2}{c|}{\textbf{Training dataset} and \textbf{Feature}}  & \multicolumn{1}{c}{\textbf{SF}} & \multicolumn{1}{c}{\textbf{Foote}} & \multicolumn{1}{c}{\textbf{CNMF}} 
     \\ \midrule 
    \multirow{5}{*}{}&   
         \multirow{1}{*}{MFCC } & \usebox{\mfccSF} & \usebox{\mfccFoote} & \usebox{\mfccCNMF} \\\cline{2-5}
        & \multirow{1}{*}{PCP} & \usebox{\pcpSF} & \usebox{\pcpFoote} & \usebox{\pcpCNMF} \\\cline{2-5}
        & \multirow{1}{*}{Tonnetz} & \usebox{\tonnetzSF} & \usebox{\tonnetzFoote} & \usebox{\tonnetzCNMF} \\\cline{2-5} 
        & \multirow{1}{*}{CQT} & \usebox{\cqtSF} & \usebox{\cqtFoote} & \usebox{\cqtCNMF} \\\cline{1-5} 
    \multirow{1}{*}{GTZAN \cite{GTZAN}}& 
        \multirow{1}{*}{\textbf{Embeddiogram}} & \usebox{\embeddioaSF} & \usebox{\embeddioaFoote} & \usebox{\embeddioaCNMF} \\\cline{1-5} 
    \multirow{1}{*}{MSD \cite{MSD}}& 
        \multirow{1}{*}{\textbf{Embeddiogram}} & \usebox{\embeddiobSF} & \usebox{\embeddiobFoote} & \usebox{\embeddiobCNMF} \\\bottomrule
    \end{tabular}
    \caption[Metric comparison: audio features and segmentation algorithms]{Comparison of precision, recall, and f-measure for different features obtained using the SF\cite{sf}, Foote \cite{Foote2000AutomaticNovelty}, and CNMF \cite{NietoCONVEXIDENTIFICATION} algorithms on the SALAMI dataset. Sliding windowed segments across the audio signal input signal is 4 seconds}\label{ta:results}
  \end{threeparttable}
  \end{adjustbox}
\end{table}

@@

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Authors [Ref], Year}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure}} \\
\midrule
\addlinespace
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{Grill2015MusicAnnotations}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\addlinespace
\cite{deepfeaturesegment}, 2019 & XXXX & CNN  & \textbf{0.535} ± 0.15 \\
\addlinespace
\cite{sf}, 2021 & XXXX & CNN  & XXX \\
\addlinespace
Ours, 2023 & Raw waveforms & CNN  & 0.288 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{\small{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. This table has been extended from \cite{Hernandez-Olivan2021MusicFeatures}.}}
\label{tab:comparison_table}
\end{table}

>>>>

The GitHub repository containing all the code needed to run the experiments can be found \href{https://github.com/oriolcolomefont/Master-Thesis.git}{HERE}.The results are displayed in the table \ref{ta:results}

@@

@@

\begin{table}
  \centering
  \begin{adjustbox}{width=\textwidth}
  \begin{threeparttable}
    \begin{tabular}{c|l|c|c|c} 
\toprule
     \multicolumn{2}{c|}{\textbf{Training dataset} and \textbf{Feature}}  & \multicolumn{1}{c}{\textbf{SF}} & \multicolumn{1}{c}{\textbf{Foote}} & \multicolumn{1}{c}{\textbf{CNMF}} 
     \\ \midrule 
    \multirow{5}{*}{}&
        \multirow{1}{*}{PCP} & \usebox{\pcpSF} & \usebox{\pcpFoote} & \usebox{\pcpCNMF} \\\cline{2-5}
        & \multirow{1}{*}{Tonnetz} & \usebox{\tonnetzSF} & \usebox{\tonnetzFoote} & \usebox{\tonnetzCNMF} \\\cline{2-5} 
        & \multirow{1}{*}{CQT} & \usebox{\cqtSF} & \usebox{\cqtFoote} & \usebox{\cqtCNMF} \\\cline{1-5} 
    \multirow{1}{*}{GTZAN \cite{GTZAN}}& 
        \multirow{1}{*}{\textbf{Embeddiogram}} & \usebox{\embeddioaSF} & \usebox{\embeddioaFoote} & \usebox{\embeddioaCNMF} \\\cline{1-5} 
    \multirow{1}{*}{MSD \cite{MSD}}& 
        \multirow{1}{*}{\textbf{Embeddiogram}} & \usebox{\embeddiobSF} & \usebox{\embeddiobFoote} & \usebox{\embeddiobCNMF} \\\bottomrule
    \end{tabular}
    \caption[Metric comparison: audio features and segmentation algorithms]{Comparison of precision, recall, and f-measure for different features using the Structural Feature (SF)\cite{sf}, Checkerboard Kernel (Foote) \cite{Foote2000AutomaticNovelty}, and Convex Non-negative Matrix Factorization (CNMF) \cite{NietoCONVEXIDENTIFICATION} algorithms on the SALAMI dataset. Sliding windowed segments across the audio signal input signal is 4 seconds long.}\label{ta:results}
  \end{threeparttable}
  \end{adjustbox}
\end{table}

@@

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Authors [Ref], Year}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure}} \\
\midrule
\addlinespace
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{Grill2015MusicAnnotations}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\addlinespace
\cite{deepfeaturesegment}, 2019 & CQT & CNN  & \textbf{0.535} ± 0.15 \\
\addlinespace
Ours, 2023 & Raw waveforms & CNN  & 0.288 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{\small{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. This table has been extended from \cite{Hernandez-Olivan2021MusicFeatures}.}}
\label{tab:comparison_table}
\end{table}
